---
title: "Communicable Disease Quarterly Report — Q3 2024"
subtitle: "Region 4 Epidemiology Unit | July 1 – September 30, 2024"
author: "Region 4 Epidemiology Data Team"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    df_print: kable
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(kableExtra)
library(lubridate)
library(here)
```

```{r load-data}
# NOTE: This report depends on the suppressed dataset.
# Run scripts 01 through 03 before knitting.
# Issue #7 (FIPS fixes) and Issue #9 (date field) must be resolved first.

cases <- read_csv(
  here("data", "cleaned", "cases_q3_2024_suppressed.csv"),
  col_types = cols(.default = col_character())
)
```

---

> **Draft status:** This report is in progress. Sections marked ⚠️ depend on 
> open issues and contain placeholder text.

---

## Summary

During Q3 2024 (July 1 – September 30), Region 4 reported a total of 
**`r nrow(cases)`** cases of notifiable communicable diseases across 
**`r n_distinct(cases$county_fips)`** counties.

```{r summary-table}
cases |>
  count(disease_group, disease, name = "Cases") |>
  arrange(disease_group, disease) |>
  kbl(caption = "Table 1. Cases by disease, Q3 2024") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  pack_rows(index = table(
    cases |> arrange(disease_group) |> pull(disease_group)
  ))
```

---

## Enteric diseases

```{r enteric-table}
cases |>
  filter(disease_group == "Enteric") |>
  count(county_name, disease, name = "n_display") |>
  pivot_wider(names_from = disease, values_from = n_display, values_fill = "0") |>
  kbl(caption = "Table 2. Enteric disease cases by county, Q3 2024") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## Vaccine-preventable diseases

```{r vax-table}
cases |>
  filter(disease_group == "Vaccine-preventable") |>
  count(county_name, disease, name = "n") |>
  kbl(caption = "Table 3. Vaccine-preventable disease cases by county, Q3 2024") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## Vector-borne diseases

```{r vectorborne-table}
cases |>
  filter(disease_group == "Vector-borne") |>
  count(county_name, disease, name = "n") |>
  kbl(caption = "Table 4. Vector-borne disease cases by county, Q3 2024") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## ⚠️ Trends over time

> This section depends on resolution of [Issue #9](../../issues/9) (date field 
> selection). Figures will be added once the primary date field is confirmed.

```{r time-trend, eval=FALSE}
# Placeholder — waiting on Issue #9

cases |>
  mutate(report_week = floor_date(as.Date(report_date), "week")) |>
  count(report_week, disease_group) |>
  ggplot(aes(x = report_week, y = n, color = disease_group)) +
  geom_line() +
  geom_point(size = 1.5) +
  labs(
    title = "Weekly case counts by disease group, Q3 2024",
    x     = NULL,
    y     = "Cases",
    color = "Disease group"
  ) +
  theme_minimal()
```

---

## Methods note

Cell suppression was applied per [`docs/suppression-policy.md`](../docs/suppression-policy.md). 
Cells with fewer than 5 cases are displayed as `<5`. 
Complementary suppression was applied where needed to prevent back-calculation.

---

## Data notes

- Source: State notifiable disease surveillance system, export date 2024-10-04
- Analysis: R `r R.version$major`.`r R.version$minor`, tidyverse `r packageVersion("tidyverse")`
- Report generated: `r Sys.time()`
